Does Anybody Really Know What Time It Is?\\ Automating the Extraction of Date Scalars

### 2016-05-03 cikm2016/cikm_dateparse.tex
some more edits and comments.

\maketitle
\begin{abstract}
Interactive visual analytics is an effective approach for the timely analysis of big data. Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow. This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning. One common scalar parsing task is extracting date and time data from string representations. Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.

>>>>

\maketitle
\begin{abstract}
Interactive visual analytics is an effective approach for the timely analysis of big data. Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow. This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning. One common scalar parsing task is extracting date and time data from string representations. Several relational database management systems (RDBMS) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.

### 2016-05-03 cikm2016/intro.tex
some more edits and comments.

\subsection{Parsing Dates}
A common version of the date parsing problem is converting columns of integers in the form \texttt{yyyyMMdd} to date scalars. \Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date. Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes). Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices). The parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.

In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (\cf Figure~\ref{fig:M2}).
Some are common, but others can be quite idiosyncratic. 
Table~\ref{tab:dateformats} shows a selection of unusual date formats found in our corpus (the meanings of the ICU formatting codes can be found in Table~\ref{tab:icuformats}). 
The first example shows a time zone in the middle of the date and a year after the time; 
the second shows a leading unmatched bracket and a colon between the date and time components; 
the third shows confusion between the seconds' decimal point and the time part delimiter; 
the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components. 
This ``long tail'' of formats means that these idiosyncrasies are the rule, not the exception, 
so a static enumeration of a small set of formats will not produce high reliability.

@@

Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted. And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.

@@

\subsection{Related Work}
\input{relatedwork}

\subsection{Contributions}

>>>>

\subsection{Parsing Dates}
A common version of the date parsing problem is converting columns of integers in the form \texttt{yyyyMMdd} to date scalars. \Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date. Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSs). Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices). The parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.

In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:M2}).
Some are common, but others can be quite idiosyncratic. 
Table~\ref{tab:dateformats} shows a selection of unusual date formats found in our corpus (the meanings of the ICU formatting codes can be found in Table~\ref{tab:icuformats}). 
The first example shows a time zone in the middle of the date and a year after the time; 
the second shows a leading unmatched bracket and a colon between the date and time components; 
the third shows confusion between the seconds' decimal point and the time part delimiter; 
the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components. 
This ``long tail'' of formats means that these idiosyncrasies are the rule, not the exception, 
so a static enumeration of a small set of formats will not produce high reliability.

@@

Several RDBMSs (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted. And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.

@@

\vidya{should this be its own section?}
\subsection{Related Work}
\input{relatedwork}

\vidya{should this be its own section?}
\subsection{Contributions}
